---
title: "The Effect of Discounting and Same-Day Shipping on Profits"
author: "Thinh Mai"
output: slidy_presentation
---

```{r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Presentation Studio Video for OfficeMate
# Date: December 12, 2019
# Author: Thinh Mai

```

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

```{r message = FALSE, echo = FALSE}

# Load packages
library(tidyverse)
library(dplyr)
library(ggplot2) 
library(gridExtra) # use to put graphs together in the same frame
library(scales)    # use to improve colors
library(janitor)  # piping function
library(kableExtra)    # clean table design
library(GGally) #used to display ggpairs
library(knitr)
library(MultinomialCI) # multinomial CI for factor variabels
library(htmlTable) # format html tables
library(Hmisc) # for corr. test of multiple variables
library(shades) # for color
library(zoo)
library(inspectdf) #load auto EDA
library(DataExplorer) #load another auto EDA

setwd("C:/Users/Mai/Downloads")
# Load data
od <- read.csv("Office_Data.csv") 

#create a column for price per unit
od$price <- od$Revenue/od$Quantity

#create a column for days difference (order processing days) between ship date and order date

od$days <- as.Date((od$Ship.Date), format="%m/%d/%Y") - as.Date(od$Order.Date, format = "%m/%d/%Y")

#create a column to show if an order receive a discount YES or NO
od$discount_eligible <- ifelse(od$Discount >0, "YES","NO")

#create a column for cost
od$cost <- od$Revenue - od$Profit

#Create a new column for Season

yq <- as.yearqtr(as.yearmon(od$Order.Date, "%m/%d/%Y") + 1/12)
      od$Season <- factor(format(yq, "%q"), levels = 1:4, 
      labels = c("winter", "spring", "summer", "fall"))
      

#create a new column for profitable
od$profitable <- ifelse(od$Profit >0, "YES", "NO")
         
         
```


How Could Advertising and Promotions Affect the Company in Sales and Revenue?
===============================================

<div style="float: right; width: 95%;">

+ Background

    - A sample of 9,994 observations on daily online orders
    - To be better effective and strategic to improve profitability
    - New variables created to help in analyzing the data set and providing recommendations
    - Cost = Revenue - Profit
    - Price = Revenue / Quantity

+ The old saying “Time is Money” holds even more truth in today’s on demand economy

    - We are in a fast-moving era
    - same day delivery service occurs when goods are collected and delivered to customers within the same day

+ Having reasonable discount offers could affect our company profit and sales targets in the long run

    - Increase value
    - Maintain brand reputation



</div>


The Power of Same-Day Shipping on Profit
===============================================

<div style="float: right; width: 95%;">

+ Same-Day shipping Generated The Most Profit on Technology

    - same day ship mode raises the most profit our company simply because it meets the customer expectations
        - People are willing to pay more to get there orders as soon as possible
    - By offering a fast and same-day delivery, we attract the customers and encourage them to place an order
        - Explains why there is a boost in our profit on Technology across all segments
       

```{r}
options(scipen = 999)
# increasing profit depending on ship method?
final_ship_toasted <- od %>%
  group_by(Segment, Category, Ship.Mode) %>%
  summarise(med_revenue = median(Profit)) %>%
  #filter(any(Category = 'Technology')) %>%
  ggplot(aes(x = factor(Category), y = med_revenue, group = Ship.Mode, fill = as.factor(Ship.Mode))) +
  geom_bar(stat = "identity",width = 0.75, position = "dodge")+
  #annotate("text",x =1, y = 27, label = "NO DISCOUNT", size = 2)+
  scale_fill_manual(values =c("#9370DB","#4B0082","#8B008B","#DDA0DD"), name = "Ship Method")+
  theme_minimal()+
  labs(x = "", y = "", title = "Same-Day Shipping On Technology \nHas a Large Impact on Profit",
       subtitle = "However there is little difference between each ship method \nin Furniture and Office Supplies") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 14), 
        plot.subtitle = element_text(hjust = 0.5, face = "italic", size = 12),
        axis.title.x = element_text(size = 10)) +
  scale_y_continuous(labels = dollar)

ship_final <- final_ship_toasted  
  #theme(legend.position = "none") 

ship_final
```

</div>


Having Discounts Has No Impact on Profit
===============================================

<div style="float: right; width: 90%;">

+ Offering Discounts Do Not Generate Enough Profit

    - TThe graph shows that even when discounts are offered,profit are not generated enough and even become negative
    - When a discount offered, it lowers the revenue and can not make up for the actual cost
    - People still makes regular purchases regardless of discounts

```{r}
options(scipen = 999)
# increasing profit with discounts?
final_disc_toasted <- od %>%
  group_by(Segment, Category, discount_eligible) %>%
  summarise(med_revenue = median(Profit)) %>%
  #filter(any(Category = 'Technology')) %>%
  ggplot(aes(x = factor(Category), y = med_revenue, group = discount_eligible, fill = as.factor(discount_eligible))) +
  geom_bar(stat = "identity",width = 0.85, position = "dodge")+ 
           #position = position_dodge())+
  annotate("text",x =0.8, y = 30, label = "NO DISCOUNT", size = 2.5)+
  scale_fill_manual(values =c("#4B0082","#8B008B"), name = "DISCOUNT")+
  theme_minimal()+
  labs(x = "", y = "", title = "Discounts on Any Categories \nDo Not Have a Significant Impact on Profit",
       subtitle = "") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 14), 
        plot.subtitle = element_text(hjust = 0.5, face = "italic", size = 12),
        axis.title.x = element_text(size = 10)) +
  scale_y_continuous(labels = dollar)

final_disc <- final_disc_toasted + theme(legend.position = "none") 

final_disc
```

</div>


Solution Proposal
===============================================

<div style="float: right; width: 95%;">

+ Shift the focus of same-day shipping on Technology to the other categories
    - Ecommerce giants like Amazon have set a high standard of fast-delivery
    - Customers then begin to expect from other retailers, including us OfficeMate
    - We need to keep up with it to meet the increased demands as well as expectations from our customers

+ OfficeMate should decrease their efforts for offering regular discounts and up for alternative solutions
     - Offer quantity discounts or bundle stock
     - Utilize the discount budget to offer same-day shipping
     - Doing these will help to attract new customers and keep customer retention

+ There is room for improvement

  - Further analysis can be done if I have more information on certain variables
  - Knowing which what the costs are for each ship mode would help to adjust the selling price to cover for the actual cost 

</div>

Contact Information
===============================================

<div style="float: right; width: 95%;">

If you have questions or would like to discuss the analysis further, please contact me at tmai@seattleu.edu. I can also provide the Technical Appendix with my complete analysis and the data if you would like more detail.


